---
title: "hw4 Q6"
author: "Lisong he"
date: "2024-03-28"
output: pdf_document
---

a)
```{r}
library(faraway)
data(sat, package = 'faraway')

sat_model <- lm(total ~ expend + salary + ratio, data=sat)
summary(sat_model)
```
b)
The intercept is estimated to be 1069.234. The coefficient of expend is 16.649, -8.823 for salary, and 6.33 for ratio. But t* for all three predictors are larger than 0.05, which are statistically insignificant. There is a suggestive negative association with teacher salary if alpha is 0.1, meaning that higher teacher salaries could potentially correspond to lower average total SAT scores.

c)
```{r}
sat_model_takers <- lm(total ~ expend + salary + ratio + takers, data=sat)
summary(sat_model_takers)
```
In this new model, the old three predictors still failed to be statistically significant, whereas the new predictor 'takers' shows a highly significant (p < 0.001) negative association with average total SAT scores.

d)
```{r}
ESR <- rstudent(sat_model_takers)
plot(ESR)
```
From the plot, observation number 44 and 48 are likely to be outliers given their large distance from y = 0.
```{r}
t <- qt(1 - 0.1 / (2*50), 50 - 5 - 1, lower.tail = TRUE)
outliers <- which(abs(ESR) > t)
print(outliers)
```
In fact, no outliers are detected under the significance level 0.1. 
