---
title: "hw4 Q3"
author: "Lisong he"
date: "2024-03-27"
output: pdf_document
---

a)
```{r}
grocery_data <- read.csv("/Users/lisonghe/Library/CloudStorage/OneDrive-JohnsHopkins/Semester 2/613 Applied Statistics and Data Analysis l/Homework/hw4/grocery-retailer.csv")
pairs(grocery_data, main = "Scatter Plot Matrix for grocery Dataset")
correlation_matrix <- cor(grocery_data)
print(correlation_matrix)
```
Based on the scatter plot matrix, there seems to be a weakly positive linear relationship between Y and X1 and a even weaker positive linear relationship between Y and X2. The plot of Y and X3 indicates that X3 is a categorical variable. Also, X1 and X2 and X3 are weakly linearly related as all proved in the correlation matrix. 

b)
```{r}
model <- lm(Y ~ X1 + X2 + X3, data = grocery_data)
summary(model)
```
Based on the regression model, the estimated regression function is 
Y = 0.0007871 * X1 -13.17e * X2 + 623.6 * X3 + 4150
Here, b1 and b3 are statistically significant whereas b2 is not.

c)
```{r}
# Residuals vs Fitted
plot(model$fitted.values, resid(model), xlab="Fitted values", ylab="Residuals", main="Residuals vs Fitted")
abline(h=0, col="red")
```
This effectively checks homoscedasticity. Here, there is no pattern detected which proves constant variance.
```{r}
# Residuals vs X1
plot(grocery_data$X1, resid(model), xlab="X1", ylab="Residuals", main="Residuals vs X1")
abline(h=0, col="red")
```

```{r}
# Residuals vs X2
plot(grocery_data$X2, resid(model), xlab="X2", ylab="Residuals", main="Residuals vs X2")
abline(h=0, col="red")
```

```{r}
# Residuals vs X3
plot(grocery_data$X3, resid(model), xlab="X3", ylab="Residuals", main="Residuals vs X3")
abline(h=0, col="red")
```
Here, the three plots of residuals against X1, X2, and X3 check the necessity of transformation of predictors. Here, all three plots show no patterns which means no transformation is needed.
```{r}
grocery_data$X1X2 <- grocery_data$X1 * grocery_data$X2
residuals <- resid(model)
if(length(residuals) == nrow(grocery_data)) {
  plot(grocery_data$X1X2, residuals, 
       xlab = "X1*X2 (Interaction Term)", 
       ylab = "Residuals", 
       main = "Plot of Residuals against X1*X2")
  abline(h = 0, col = "red")
} else {
  stop("The lengths of 'residuals' and 'grocery_data$X1X2' do not match.")
}
```
Plotting residuals against interaction terms checks the presence of the interaction term
```{r}
# Normal Q-Q plot
qqnorm(resid(model), main="Normal Q-Q Plot")
qqline(resid(model), col="red")
```
This testifies the normality of residuals. At two ends of the distribution, the dots deviate from the normal line which stands for longer tails than a normal curve. The residual is roughly normal.

d)
```{r}
model_2 <- lm(Y ~ X1 + X2, data=grocery_data)
summary(model_2)
```

e)
```{r}
library(car)
avPlots(model_2)
```
Added-Variable plots plot residuals of Y after regressing on all predictors except Xi against that of Xi. It is essentially the plot of part of Y unexplained by other predictors against part of Xi unexplained by other predictors. Both plots show non-zero slopes, indicating that X1 explains Y.

f)
```{r}
model_Y_on_X1 <- lm(Y ~ X1, data=grocery_data)
model_X2_on_X1 <- lm(X2 ~ X1, data=grocery_data)
model_resids <- lm(resid(model_Y_on_X1) ~ resid(model_X2_on_X1))
summary(model_resids)
```

