---
title: "q3"
author: "Lisong He"
date: "2024-04-10"
output: pdf_document
---
a)
```{r}
data <- read.csv('grocery-retailer.csv')
model <- lm(Y ~ X1 + X2 + X3, data=data)
hat_values <- hatvalues(model)
print(hat_values)
```
```{r}
outlying_X <- which(hat_values > 2*4/52)
outlying_X
```
b)
```{r}
plot(data$X1, data$X2, main="Scatter plot of X2 against X1")
X_new <- cbind(1, 300, 7.2, 0) # Assuming your model has an intercept
predict(model, newdata=data.frame(X1=300, X2=7.2, X3=0))

```
Visually at X1 = 300000 and x2 = 7.2, the point is in the center of the data so no extrapolation.
```{r}
X <- model.matrix(model, data)
XTX <- t(X) %*% X
XTX_inv <- solve(XTX)
X_new <- matrix(c(1, 300, 7.2, 0), nrow = 1)
lev <- X_new %*% XTX_inv %*% t(X_new)
print(paste('Hnew,new is ',lev, 'for this measurement.'))
print(paste('Range of hat values is [', min(hat_values), ',', max(hat_values), ']'))
```
Apparently, Hnew,new exceeds the bounds so it is a hidden extrapolation, which contradicts my previous observation.

c)
Looking at DFFITS:
```{r}
cases <- c(16, 22, 43, 48, 10, 32, 38, 40)
for (case in cases) {
  cat("Case", case, "has DFFITS of ", dffits(model)[case], "\n")
  if (abs(dffits(model)[case]) > 2 * sqrt(4/52)) {
    cat("Case", case, "is an influential observation on Y in terms of DFFITS", "\n")
  } else {
    cat("Case", case, "is not a influential observatio on Y in terms of DFFITS", "\n")
  }
  cat("\n")
}
```
Looking at cook's distance:
```{r}
cases <- c(16, 22, 43, 48, 10, 32, 38, 40)
for (case in cases) {
  cook <- cooks.distance(model)[case]
  cat("Case", case, "has Cook's distance of", cook, "\n")
  if (cook <= 0.2) {
    cat("Case", case, "has no influence in terms of Cook's distance\n")
  } else if (cook > 0.2 && cook <= 0.5) {
    cat("Case", case, "has moderate influence in terms of Cook's distance\n")
  } else if (cook > 0.5) {
    cat("Case", case, "has major influence in terms of Cook's distance\n")
  }
  cat("\n")
}
```
Looking at DFBETAS:
```{r}
cases <- c(16, 22, 43, 48, 10, 32, 38, 40)
for (case in cases) {
  cat("Case", case, "has DFBETAS of ", dfbetas(model)[case], "\n")
  if (abs(dfbetas(model)[case]) > 2/sqrt(52)) {
    cat("Case", case, "is an influential observation on Y in terms of DFBETAS", "\n")
  } else {
    cat("Case", case, "is not a influential observatio on Y in terms of DFBETAS", "\n")
  }
  cat("\n")
}
```
d)
```{r}
plot(cooks.distance(model), type="h", main="Index plot of Cook's distance")
```
According to the plot, none of the cases has cook's distance larger than 0.1 so they are all considered non-influential.

e)
```{r}
# Scatter plot matrix
pairs(data[, c("X1", "X2", "X3")])
```
```{r}
# Correlation matrix
cor(data[, c("X1", "X2", "X3")])
```
The scatter plots and correlation coefficients indicate that there are weak linear relationships between the variables. The binary nature of X3 results in distinct groupings rather than a traditional scatter pattern, which could influence the low correlation coefficients with X1 and X2.

f)
```{r}
library(car)
vif(model)
```

