---
title: "hw5 q1"
author: "Lisong he"
date: "2024-04-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.
a)
```{r}
library('TSA')
series = arima.sim(n=48,list(ar=0.7))
theoretical_acf_lag1 <- 0.7
theoretical_acf_lag5 <- 0.7^5
print(paste('Theoretical autocorrelation at lag 1 is', theoretical_acf_lag1))
print(paste('Theoretical autocorrelation at lag 5 is', theoretical_acf_lag5))
```
b)
```{r}
sample_acf_lag1 <- acf(series, plot = FALSE)$acf[1]
sample_acf_lag5 <- acf(series, plot = FALSE)$acf[5]
var_lag1 <- (1-0.7^2)/48
var_lag5 <- (1+0.7^2)/(48*(1-0.7^2))
z_score_lag1 <- (sample_acf_lag1 - theoretical_acf_lag1) / sqrt(var_lag1)
z_score_lag5 <- (sample_acf_lag5 - theoretical_acf_lag5) / sqrt(var_lag5)
print(paste('Z-score of sample autocorrelation at lag 1 is', z_score_lag1))
print(paste('Z-score of sample autocorrelation at lag 5 is', z_score_lag5))
```
Here, both sample acfs have z-scores below the absolute value of 1.96, suggesting that they are not significantly different from the theoretical acfs.

c)
```{r}
series1 <- arima.sim(n=48,list(ar=0.7))
sample1_acf_lag1 <- acf(series1, plot = FALSE)$acf[1]
sample1_acf_lag5 <- acf(series1, plot = FALSE)$acf[5]
z_score1_lag1 <- (sample1_acf_lag1 - theoretical_acf_lag1) / sqrt(var_lag1)
z_score1_lag5 <- (sample1_acf_lag5 - theoretical_acf_lag5) / sqrt(var_lag5)
print(paste('Z-score of sample autocorrelation at lag 1 is', z_score1_lag1))
print(paste('Z-score of sample autocorrelation at lag 5 is', z_score1_lag5))
```
In this new simulation, the z-scores are larger in scale than the previous one, but they still lie within the 95% confidence interval, suggesting that sample acfs are not significantly different from theoretical values.

d)
```{r}
r1_samples <- numeric(10000)
r5_samples <- numeric(10000)
for(i in 1:10000) {
  sim_data <- arima.sim(n = 48, list(ar = 0.7))
  acfs <- acf(sim_data, lag.max = 5, plot = FALSE)$acf
  r1_samples[i] <- acfs[1]
  r5_samples[i] <- acfs[5]
}
hist(r1_samples, main="Histogram of r1", xlab="Autocorrelation at lag 1")
hist(r5_samples, main="Histogram of r5", xlab="Autocorrelation at lag 5")
sample_var_r1 <- var(r1_samples)
sample_var_r5 <- var(r5_samples)
large_sample_var_r1 <- var_lag1
large_sample_var_r5 <- var_lag5

print(paste('Sample variance of r1 is', sample_var_r1))
print(paste('Large sample variance of r1 is', large_sample_var_r1))
print(paste('Sample variance of r5 is', sample_var_r5))
print(paste('Large sample variance of r5 is', large_sample_var_r5))
```
Here, the sample and theoretical variances are very close to each other for lag1 and relatively close at lag 5.
